An investigation into using scagnostics to find interesting projections of high-dimensional data

### 2018-10-17 scagnostics.Rmd
More text.

We now

@@

## Example 2
The above example is using only a reduced set of the free parameters relevant to the physical interpretation, and we already know the expected result by construction. In general many additional nuissance parameters will also enter the fit, see e.g. @Smith:2016qas. We use a simulation study \textcolor{red}{simulating a binary black hole (BBH) merger event...need more info maybe, no reference but Rory offered to write short description} including the full set of nuissance parameters as a test case of larger parameter space in which we aim to detect structure. The dataset is summarised in \ref{fig:bbhSimulation} where we have removed a subset of less informative parameters to obtain a clear picture. Interesting correlations can be seen between the parameters "ra", "dec" and "time", where the first two describe the position of the BBH event in the sky, and the "time" being the merging time (in GPS units). Note in particular the relation between "dec" and "time" is not functional, which motivates us to use the TIC index for this example. Some other interesting patterns can also be identified.
\textcolor{red}{Question for Rory: is dependence between dec, time, ra built into prior distribution?}
Show three results:
 - given all variables, including nuissance, recover dec, time, ra from guided tour (TIC, splines, tried 1-convex but maybe another example is better? skinny?)
 - removing those three, do we find interesting views?
 - using smaller subset, find improved description when using more than 2 parameters

@@

>>>>

To further investigate the dataset we remove the $L1$ parameter and again use the spline index to find functional dependence between the remaining four parameters. For this example, because of the strong correlation between the various parameters, we recover the dependence of $L2$ on the mass and radius of the lighter neutron star, despite having removed its mass from the parameter set. In this setting $L2$ is best described as a function of all 3 parameters still considered, see Figure \ref{fig:nseRemL1}.

@@

## Example 2
The above example is using only a reduced set of the free parameters relevant to the physical interpretation, and we already know the expected result by construction. In general many additional nuissance parameters will also enter the fit, see e.g. @Smith:2016qas. We use a simulation study \textcolor{red}{simulating a binary black hole (BBH) merger event...need more info maybe, no reference but Rory offered to write short description} including the full set of nuissance parameters as a test case of larger parameter space in which we aim to detect structure. The dataset is summarised in \ref{fig:bbhSimulation} where we have removed a subset of less informative parameters to obtain a clear picture. Interesting correlations can be seen between the parameters "ra", "dec" and "time", where the first two describe the position of the BBH event in the sky, and the "time" being the merging time (in GPS units). Note in particular the relation between "dec" and "time" is not functional, which motivates us to use the TIC index for this example (drawback is however the run time, this data sample contains 10k points, making optimisation of the MINE index functions very inefficient). Some other interesting patterns can also be identified. (These patterns are representative of the likelihood function, since most priors are flat, or build such to capture growth with volume rahter than distance. This is still efficient thanks to special sampling techniques that allow jumps in parameter space.)

As a first test we use the full dataset (with exception of $m2$ for the same reasons as stated in the first example), and use different index functions. Here we expect to recover some relation between the correlated variables dec, time and ra. We compare the final view using three different metrics in Fig. \ref{fig:bbhGuided}, showing the views with (local?) maximum of the TIC index (left) spines2d index (middle) and for comparison maximum of 1-convex (right). Indeed the TIC and spline index identify dec, time and ra parameters and show views of the 2-d band that is embedded in three dimensions \textcolor{red}{maybe having tour of those three dimensions would be helpful here}. Notably, the exact views differ based on the index function selected, optimising TIC is showing us the band "from the side", resulting in clearly non-functional dependence. On the other hand, the index based on the scagnostics member convex is too noisy to find any views of interest, and the final  view is essentially equal to the starting plane in this case. \textcolor{red}{different scagnostics member may be better suited, but I expect the result will be the same. might still be useful to try skinny instead of 1-convex?}.

@@

We next follow a similar strategy as above and remove the variables with the strongest association from the dataset to understand if additional structure is encoded in the distribution. Removing thus time, dec and ra variable from the input, we optimise again the TIC, splines2d and 1-convex index functions. The results are shown in Fig. \ref{fig:bbhGuided2}. In this case the TIC index (left) identifies a view presenting seemingly linear correlation, originating however from one well constrained direction orthogonal to the main distribution of points. This direction is defined by the chi\_tot and chi\_p variables, for which we can read off the strong correlation in Fig. \ref{fig:bbhSimulation}. This direction is also found by the spline index (middle view), which in addition identified the bimodal structure in the data stemming from the degeneracy between two plausible values for theta\_jn. (Interestingly the hole index in two dimensions does not identify this view, but is expected to find it in a 1-d guided tour.) Finally, as shown on the right, 1-convex does not identify any views of interest here.
